---
title: "slides-bootstrap"
author: "Ningning Long"
date: "November 30, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Bootstrap Procedure 

Bootstrap for the factor scores is for estimating the stability of the compromise factor scores.

First, we sample from the table index. If the dataset is combined with k tables, we sample from 1 to k for k times, which becomes our first sample set. Second, we compute the partial factor scores for our sample sets, and we denote it as $F_1,\dots,F_k$, and denote F$_1^*$=$\frac{1}{k}\sum_{i=1}^k F_i$ as the bootstrapped estimate of the factor scores from the first sample.

Repeat step 1 and 2 for large number of times (e.g., L=1000), so we can get F$_1^*$,F$_2^*$,$\dots$,F$_L^*$.

Then we can caculate the mean and variance of bootstrap samples:
  $$\bar{F}^*= \frac{1}{L}\sum_{l=1}^L F_l^*$$
  $$\hat{\sigma}_{F^*}^2=\frac{1}{L}(\sum_{l=1}^{L}( F_l^*-\bar{F}^*)\cdot(F_l^*-\bar{F}^*))$$

Finally, we can get the bootstrap ratio by $\frac{\bar{F}^*}{\hat{\sigma}_{F^*}}$. The bootstrap ratio gives information of which observations most reliably contribute to a component.

  

##Bootstrap Method in the MFA Package
The bootstrap is a method for mfa objects

usage: bootstrap(mfa,value)

inputs: a mfa object and the number of times of bootstrap

outputs: a list, containing the mean, standard deviation and bootstrap ratio; Also, it creates plot of bootstrap ratio for each dimension.


          
##Example of Bootstrap
```{r}
#use MFA package
library(MFA)

#get the dataset
wines <- read.csv("https://raw.githubusercontent.com/ucb-stat243/stat243-fall-2016/master/problem-sets/final-project/data/wines.csv", stringsAsFactors = FALSE)

#According to the dataset, creating a list of vectors indicating the sets of variables
sets <- list(2:7, 8:13, 14:19, 20:24, 25:30, 31:35, 36:39, 40:45, 46:50, 51:54)

#The scaling vector
scaling_vec <- apply(subset(wines, select = unlist(sets)), 2, function(x) sqrt(sum((x - mean(x))^2)))

#create the mfa object 
mymfa <- mfa(wines, sets, ncomps = 2, T, scaling_vec)

#do boostrap 1000 times with a mfa object
bootstrap(mymfa, 1000)
```

          